---
title: "Finance Quantitative"
subtitle: "Modèlisation du Smile"
author: "Le groupe de travail"
date: "Version: `r format(Sys.Date(), '%d %b %Y')`"
output:
  pdf_document:
    keep_tex: false
    fig_caption: yes
    latex_engine: pdflatex
geometry: margin=1in
header-includes:
  - \usepackage[utf8]{inputenc}
  - \usepackage{float}
  - \usepackage{eurosym}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r load-libraries, include=FALSE, echo=FALSE}
library(kableExtra)
library(lubridate)
library(BondValuation)
library(lpSolve)
get.src.folder <- function() {
path.expand("../GP/src")
}

get.data.folder <- function() {
  path.expand("../GP/data")
}

source(file.path(get.src.folder(), 'utils.R'))
source(file.path(get.src.folder(), 'FileUtils.R'))
```

In this problem set, you will use the following functions:

GBSPrice: Price of a vanilla option:
\[
P = f(\text{PutCall}, S, K, T, r, b, \sigma)
\]

where:
\begin{description}
\item[$\text{PutCall}$] 'c' for a call, 'p' for a put
\item[$b$] cost of carry: ridk free rate $r$ less dividend yield $d$
\item[$r$] risk-free rate
\end{description}

```{r GBSPrice, eval=T, echo=T}
GBSPrice <- function(PutCall, S, K, T, r, b, sigma) {
  d1 <- (log(S/K) + (b+sigma^2/2)*T)/(sigma*sqrt(T))
  d2 <- d1 - sigma*sqrt(T)

  if(PutCall == 'c')  
    px <- S*exp((b-r)*T)*pnorm(d1) - K*exp(-r*T)*pnorm(d2)
  else
    px <- K*exp(-r*T)*pnorm(-d2) - S*exp((b-r)*T)*pnorm(-d1)

px
}
```

GBSVega: Vega ($\frac{\partial P}{\partial \sigma}$) of a Vanilla option:

```{r GBSVega, echo=T}
GBSVega <- function(PutCall, S, K, T, r, b, sigma) {
  d1 <- (log(S/K) + (b+sigma^2/2)*T)/(sigma*sqrt(T))
  S*exp((b-r)*T) * dnorm(d1)
}
```

# Data

The spot is $S=110$. We observe the following prices for calls and puts on an asset paying a continuous dividend.

```{r data-set, echo=F, eval=T}
df <- read.csv(file.path(get.data.folder(), "call-put-prices.csv"))
kable(df[, c("Strike", "Call", "Put")], booktabs=TRUE, digits=c(0,3,3)) %>%
  kable_styling()
```


# Questions

## Dividend yield and risk-free rate

Using the Call-Put parity, estimate by linear regression the implied risk-free rate ($r$) and dividend yield ($d$).

- Using the functions above, write a function that computes the implied volatility of a Vanilla option. The function should have the following signature:

```{r}
ImpliedVol <- function(p, TypeFlag, S, X, Time, r, b, sigma=NULL, maxiter=500, tol=1.e-5) {
}
```

where:
\begin{description}
\item[$p$] price of the option
\item[$sigma$] an optional initial value for the volatility
\item[$maxiter$] an optional maximum number of iterations
\item[$tol$] an optional tolerance for the error $|g(\sigma)|$.
\end{description}

- Test the accuracy of your procedure on options that are deep in the money and deep out of the money, and report the results of your tests.
- Compute the implied volatility of the calls and puts in the data set.
- Fit a quadratic function to the call and put implied volatilities (one function for the calls, one for the puts), and plot actual vs. fitted data. Interpret the results.


## Breeden-Litzenberger formula

Compute the implied density of $S_T$ using the Breeden-Litzenberger formula. Estimate 
\[
\frac{\partial^2 f}{\partial K^2}
\]
by finite difference. Remember that now $\sigma$ is a function of strike. Plot the implied distribution and compare to the distribution implicit in the standard Black-Scholes model. Interpret your observations.

## Shimko's Model

Compute the implied distribution of $S_T$ using Shimko's model and the quadratic smile function estimated above.
Plot this distribution and compare with the result of the Breeden-Litzenberger formula. Interpret your observations.


## Pricing a digital call

Recall that a digital call with strike $K$ pays one euro if $S_T \geq K$, and nothing otherwise.

Using the implied density computed above, compute the price of a digital call by numerical integration.

Perform this calculation for strikes ranging from 80 to 140. Compare with the price obtained using a log-normal distribution for $S_T$. Interpret your observations.
