---
title: "Finance Quantitative"
subtitle: 'Pricing Vanna-Volga'
author: Le groupe de travail
date: "Version: `r format(Sys.Date(), '%d %b %Y')`"
output:
  pdf_document:
    keep_tex: false
    fig_caption: yes
    latex_engine: pdflatex
  word_document: default
geometry: margin=1in
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{amsmath}
- \usepackage{amsfonts}
- \usepackage{amssymb}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(kableExtra)
```
The purpose of this problem set is to explore the Vanna-Volga pricing model.

# Getting started

- Using the Rmetrics library (fOptions), verify that you know how to compute the price and the "greeks" of a vanilla option.
- Identify or write a robust function to compute the implied volatility, given the price of a vanilla option.

# Questions

## Volatility Interpolation

Given the implied volatility at three strikes, we will use the Vanna-Volga pricing method to interpolate the
volatility curve. Assume $r=0, b=0, T=1, \mbox{Spot} = 100$. The Black-Scholes volatility
for three strikes is given below.

```{r, echo=FALSE}
# Benchmark data: (strike, volatility)
VolData <- list(c(80, .32), c(100, .30), c(120, .315))
tmp <- as.data.frame(VolData)
row.names(tmp) <- c("Strike", "Volatility")
colnames(tmp) <- NULL
kable(tmp, booktabs=T) %>%
  kable_styling(position = "center", latex_options = "HOLD_position")
```

1. Write utility functions to compute the risk indicators Vega, Vanna and Volga,
 all by finite difference.
 
2. Compute vectors of vega, vanna, volga for the three benchmark instruments

3. Choose a new strike for which you want to compute the implied volatility.

4. Compute the risk indicators for a call option struck at that strike. 
    
5. Compute the Vanna-Volga price adjustment and the corresponding implied volatility.
    
6. Wrap the above logic in a function in order to
interpolate/extrapolate the volatility curve from $K=70$ to $K=130$. Draw a plot of the interpolated volatility curve.

## Pricing a digital call

Recall that a digital call with strike $K$ pays one euro if $S_T \geq K$, and nothing otherwise.

Using the same logic as in the previous question, price a digital call, maturity $T=1$, struck at $K=105$. 
